Grow with the flow: faster growth is associated with more variable precipitation in a perennial herb

### 2016-10-07 ms/ms.Rnw
Worked on ms

%* put in biomass allocation cline
% TO DO:
% 1. Is water regime on figure 1 correct?
% 2. Finalize Table 2
% 3. Do anything with biomass data?
% 4. Citations for latitudinal clines
	% cases where clinal variation has 'obvious' selective source:
	% Antonovics and Bradshaw discuss steepness of clines and strength of selection over 10's of meters off mines in Anthoxanthum
	% Caicedo et al 2004?
	% clines in phenotypic traits reviewed by Hedrick 2006, Schemske and Bier 2007
% 5. Citations for GxE as signature of local adaptation
% 6. Fix figure 2
% 7. Supplemental Figure and Table captions
% 8. Supplemental figure showing agreement between assigned to climatic variables from point estimated and spatially averaged analyses

@@

# Read in spatially-averaged (100 km buffer) for 16 focal populations
load(paste(pathDat, "/focClimSA.RData", sep = ""))

@@

with(exptlDes, text(min((which(PhotoMeas)) + max(which(PhotoMeas))) / 2, 5, 
	labels = "Photo", col = "grey", font = 3, cex = 0.9))
with(exptlDes, rect(min(which(PhotoMeas)), 2, max(which(PhotoMeas)), 4, 
	border = "white", col = "grey"))
#with(exptlDes, text(min((which(PhotoMeas)) + max(which(PhotoMeas))) / 2, 3, 
#	labels = "Photo", col = "white", font = 3, cex = 0.9))

@@

# with(exptlDes, points(DayN, WaterFreq_dry, type = "l", lwd = 12, 
	# col = alpha("white", 0.5)))
with(exptlDes, points(DayN, WaterFreq_dry, type = "l", lwd = 3, col = "black"))

@@

@@

  # Mixed model ANOVA with family as random factor fit using step down procedure
	master$Chamber <- substr(master$Block, 1, 3)
	mm <- lmer(LER_AbsGrowth ~ AvgGermDay + PopID * TempTrt * WaterTrt + 
						 	(1|Family) + (1|Block), data = master)
	fitLER <- step(mm, reduce.random = F) # includes pop, temp, h2o, population x water
	# fitLER <- step(mm, ddf = "Kenward-Roger", reduce.random = F) # same result using Satterthwaite ddf

@@

  mm <- lmer(SER_AbsGrowth ~ AvgGermDay + PopID * TempTrt * WaterTrt + 
		(1|Family) + (1|Block), data = master)
	fitSER <- step(mm, reduce.random = F) #includes pop, temp
@

\paragraph{Photosynthesis}
During the week of 10 to 16 June (57 to 63 days after sowing), we measured daytime photosynthetic rate and stomatal conductance on a subset of \Sexpr{length(which(!is.na(master$Photo)))} plants evenly spread between treatments and families within populations. The youngest, fully-expanded leaf \cdm{this is what I did...not sure exactly which node, as I didn't record that when I did measurements} acclimated for 3 minutes to reach steady state in a 6 cm$^2$ chamber of a LI-COR 6400XT Portable Photosynthesis System (LI-COR Biosciences, Lincoln, Nebraska). All measurements were made at ambient light (400 $\mu$mol m$^{-2}$ s$^{-1}$), temperature, and moderate relative humidity. During this period, we suspended normal day-to-day temperature fluctuations and set daytime temperatures to its average for that period (Cool: 26.5$\degree$; Hot: 36.1$\degree$ so that all plants within a temperature level were measured under the same conditions.

@@

  fitMort <- glmer(DiedFromStress ~ PopID * TempTrt * WaterTrt + (1|Family) +
  								 (1|Block), control = glmerControl(optimizer = "bobyqa"),
  								 data = subset(master, !is.na(master$DiedFromStress)), 
  								 family = "binomial")
  # Anova(fitMort, type = 2)
@

@@

  # plot axes could be range maps with areas highlighted (x-axis) and short to tall plants (y-axis)
  #plot(focClim$Lat, traitPC$x[, "PC1"])
  #cor.test(focClim$Lat, traitPC$x[, "PC1"])
  #cor.test(popMeans[, 4], traitPC$x[, "PC1"])
  #plot(focClim$bio16_cva, traitPC$x[, "PC1"])
@

@@

	# Load climVars
	load(paste(pathDat, "/ExploratoryCimateVariables.RData", sep = ""))

@@

  # Variable selection using random forests (VSURF)
  linmod.PC1 <- as.formula(paste("PC1 ~ ", paste(climVars, collapse = " + "), sep = ""))
  set.seed(69648)
  # vsurfPC1 <- VSURF(linmod.PC1, data = focClim[, c("PC1", climVars)])
  vsurfPC1 <- VSURF(focClim[, climVars], focClim[, "PC1"])
  # vsurfPC1_SA <- VSURF(linmod.PC1, data = focClimSA[, c("PC1", climVars)])
  vsurfPC1_SA <- VSURF(focClimSA[, climVars], focClimSA[, "PC1"])

##### WORKING ON STUFF HERE #####
#vsurfPC1_SA$imp.mean.dec
#x <- match(vsurfPC1$imp.mean.dec.ind, vsurfPC1_SA$imp.mean.dec.ind)
#plot(vsurfPC1$imp.mean.dec, vsurfPC1_SA$imp.mean.dec[x], type = "n")
#text(vsurfPC1$imp.mean.dec, vsurfPC1_SA$imp.mean.dec[x], 
#		 labels = climVars[vsurfPC1$imp.mean.dec.ind])
#cor.test(vsurfPC1$imp.mean.dec, vsurfPC1_SA$imp.mean.dec[x])

@@

  # polygon of confidence intervals
  x <- seq(min(focClim$bio16_cva), max(focClim$bio16_cva), l = 1e3)
  u <- predict(fitBio16cva, new = data.frame(bio16_cva = x))
  s <- sqrt(sapply(x, function(X) t(c(1, X)) %*% vcov(fitBio16cva) %*% c(1, X)))
  confInt <- data.frame(lowerCI = qnorm(0.025, u, s), upperCI = qnorm(0.975, u, s))
  polygon(c(x, rev(x)), c(confInt$lowerCI, rev(confInt$upperCI)), 
    col = "grey80", border = NA)
  lines(x, confInt$lowerCI, lwd = 2, lty = 2)
  lines(x, confInt$upperCI, lwd = 2, lty = 2)
  lines(range(focClim$bio16_cva), range(predict(fitBio16cva)), lwd = 2)
  points(focClim$bio16_cva, focClim$PC1, pch = 21, col = "black", bg = primShade0, 
    cex = 1.5, lwd = 2)
  dev.off()

@@

@@

	\subsection{Structure}

	Leaf traits method: \\
	Day 1: \\
	- select focal leaf (e.g. second or third leaf); \\
	- place in H20, dark, cool overnight. \\
	Day 2: \\
	- wet weight; \\
	- scan; \\
	- place in drying oven. \\

		\begin{enumerate}
			\item{Stomatal density, guard cell length, ratio from leaf opposite focal leaf (above}
			\item{Leaf size/shape (from scans)}
			\item{Leaf thickness/SLA}
			\item{Colorimetry (from scans, spec)}
			\item{stem anthocyanin (qualitative)}
		\end{enumerate}

	OPEN QUESTIONS: \\
	save wet samples in preservative (e.g. for leaf venation?) \\
	chlorophyll content? \\
	branching architecture (anything in plant protocols?)

	\subsection{Function}

		\subsubsection{during experiment}

		\begin{enumerate}
			\item{Instantaneous photosynthetic rate, $g_\text{s}$, water-use efficiency.}
				\subitem{3 min per plant per LICOR}
				\subitem{Do as many plants as possible in 1-2 weeks}
				\subitem{Before and after drought?}
			\item{Respiration}
			\item{Whole plant water-use (gravimetric)}
				\subitem{}
			\item{Leaf temperature (subset)}
			\item{Temperature response curve (subset)}
				\subitem{see protocol here: http://prometheuswiki.publish.csiro.au/tiki-index.php?page=Temperature+response+of+photosynthesis+using+a+Li-Cor+6400}
			\item{Light response curve}
				\subitem{Sampling? Only in cool, wet?}
				\subitem{At what temperature?}

		\end{enumerate}

		\subsubsection{after experiment}

		\begin{enumerate}
			\item{CN content}
			\item{carbon isotope discrimination}
		\end{enumerate}

	\subsection{Development time and Life history}

	\begin{enumerate}
		\item{Flowering time: census daily}
		\item{Starch accumulation: collect (subset) of plants}
		\item{Rhizome formation: collect on subset of plants harvested for dry mass}
	\end{enumerate}

	\subsection{Floral (2nd/3rd/etc flower?)}

	\begin{enumerate}
		\item{Flower size}
		\item{Herkogamy}
	\end{enumerate}

\section{Performance}

	Possibilities: \\
	1. (sub-)weekly census of plant height. $\approx$ 0.5 min per plant, so allow 1 day \\

	$$RGR_\text{height} = \frac{\text{ln} H_2 - \text{ln} H_1}{t_2 - t_1}$$

	2. early development (pre-treatment) leaf-area growth with camera \\
	3. harvest dry biomass before and after treatments (get whole plant leaf area using scanner at same time to ground truth camera)

	$$RGR_\text{mass} = \frac{\text{ln} M_2 - \text{ln} M_1}{t_2 - t_1}$$

\section{Determining ecological gradients that matter}

Performance tradeoffs
Isolation by ecology

\section{Tests of divergent selection}

	\subsection{Trait-environment correlation}

	We will account for relatedness by using genetic distance matrix in linear models.

	\subsection{$P_\text{ST}$ -- $F_\text{ST}$}

	Contrast trait classes with Fst distribution

	\subsection{Isolation by ecology}

	Is \textbf{G} better predicted by \textbf{E} than geographic distance? Use Gideon's method?

	\subsection{Performance tradeoffs between treatments}

\section{Ecological strategy, physiological mechanism, and tradeoff}

	\subsection{Drought escape}
		Pattern: \\
			Flowering time shorter in drier/S populations \\
			More between population variation in flowering time more variable than predicted by Fst \\
			... \\

	\subsection{Drought avoidance}
		Pattern: \\
			Maybe this should really be dehydration avoidance? \\
			Populations from drier/S avoid dehydration by limiting water loss; more variable than predicted by Fst \\

			Mechanism \\
			closing stomata earlier during soil drying

		Mechanisms: \\
			lower water-use \\

	\subsection{Drought tolerance}
		Pattern: \\
			drier/S populations have greater tolerance to low tissue water content

		Mechanisms: \\
			greater WUE \\
			ability to photosynthesize during drought \\

\section{Extension: incipient speciation as a byproduct or divergence?}

Collect pollen to measure viability.

(this also goes into the goal of trying to figure out ecological selective agent driving trait variation). Why does latitude predict trait variation, often better than any of the climatic variables that vary with latitude? Some hypotheses:
\begin{itemize}
  \item{H1: Latitude encompasses multivariate climatic dimension.}
  \item{H2: Latitude is substitute for average selective environment. i.e. populations are connected by gene flow, such that alleles at a given site must do well there and in surrounding areas.}
\end{itemize}

Predictions:
\begin{itemize}
  \item{P1: Climatic axis in high dimensional climate space should correlate with latitude (across card sites especially, not just any site [using pseudoabsences]). Focal site position in high-dimensional climatic space should correlated with traits as well or better than latitude.}
  \item{P2: Climate (averaged over nearby sites) varies with latitude and the nearby average climate predicts trait values as well or better than latitude.}
  \item{P1+2: These are not mutually exclusive and both could be imortant. i.e. average selective environment in high-dimensional climate space predicts trait values as well or better than latitude.}
\end{itemize}

>>>>

%* put in biomass allocation cline
% TO DO:
% 1. Is water regime on figure 1 correct?
% 2. Finalize Table 2
% 3. Do anything with biomass data?
% 4. Citations for latitudinal clines
	% cases where clinal variation has 'obvious' selective source:
	% Antonovics and Bradshaw discuss steepness of clines and strength of selection over 10's of meters off mines in Anthoxanthum
	% Caicedo et al 2004?
	% clines in phenotypic traits reviewed by Hedrick 2006, Schemske and Bier 2007
% 5. Citations for GxE as signature of local adaptation
% 6. Fix figure 2
% 7. Supplemental Figure and Table captions
% 8. Supplemental figure showing agreement between assigned to climatic variables from point estimated and spatially averaged analyses
% 9. new spatially averaged analysis
% 10. new analysis with updated ENM
% 11. mention using LER_AbsGrowth as predictor in fitSER
% 12. For Figure_PC1vLat: plot axes could be range maps with areas highlighted (x-axis) and short to tall plants (y-axis)

@@

# Read in spatially-averaged (62 km buffer) for 16 focal populations
# *** still need to make this filer ***
load(paste(pathDat, "/focClimSA_62km.RData", sep = ""))

@@

with(exptlDes, text(min((which(PhotoMeas)) + max(which(PhotoMeas))) / 2, 5, 
	labels = "Photo", col = "grey", font = 3, cex = 0.9))
with(exptlDes, rect(min(which(PhotoMeas)), 2, max(which(PhotoMeas)), 4, 
	border = "white", col = "grey"))

@@

with(exptlDes, points(DayN, WaterFreq_dry, type = "l", lwd = 3, col = "black"))

@@

	# Note that growth data used in the mixed model are from the master datasheet.
	# The code below shows how growth rates were calculated.

@@

  # Mixed model ANOVA with family as random factor fit using step down procedure
	mm <- lmer(LER_AbsGrowth ~ AvgGermDay + PopID * TempTrt * WaterTrt + 
						 	(1|Family) + (1|Block), data = master)
	fitLER <- step(mm, reduce.random = F) # includes pop, temp, h2o, population x water

@@

  mm <- lmer(SER_AbsGrowth ~ AvgGermDay + LER_AbsGrowth + 
               PopID * TempTrt * WaterTrt + (1|Family) + (1|Block), 
             data = master)
	fitSER <- step(mm, reduce.random = F) #includes LER_AbsGrowth, pop, temp, h2o, and temp x h2o

\paragraph{Photosynthesis}
During the week of 10 to 16 June (57 to 63 days after sowing), we measured daytime photosynthetic rate and stomatal conductance on a subset of \Sexpr{length(which(!is.na(master$Photo)))} plants evenly spread between treatments and families within populations. The youngest, fully-expanded leaf \cdm{this is what I did...not sure exactly which node, as I didn't record that when I did measurements} acclimated for 3 minutes to reach steady state in a 6 cm$^2$ chamber of a LI-COR 6400XT Portable Photosynthesis System (LI-COR Biosciences, Lincoln, Nebraska). All measurements were made at ambient light (400 $\mu$mol m$^{-2}$ s$^{-1}$), atmospheric CO$_2$ (400 ppm), temperature, and moderate relative humidity. During this period, we suspended normal day-to-day temperature fluctuations and set daytime temperatures to its average for that period (Cool: 26.5$\degree$; Hot: 36.1$\degree$ so that all plants within a temperature level were measured under the same conditions.

@@

  fitMort <- glmer(DiedFromStress ~ PopID * TempTrt * WaterTrt + (1|Family) +
  								 (1|Block), control = glmerControl(optimizer = "bobyqa"),
  								 data = subset(master, !is.na(master$DiedFromStress)), 
  								 family = "binomial")
  # Anova(fitMort, type = 2)

@@

@@

	# Load climVars
	load(paste(pathDat, "/ExploratoryClimateVariables.RData", sep = ""))

@@

  # Variable selection using random forests (VSURF)
	# Climate variables that predict PC1 (point and spatially averaged)
  set.seed(69648)
  vsurfPC1 <- VSURF(focClim[, climVars], focClim[, "PC1"])
  vsurfPC1_SA <- VSURF(focClimSA[, climVars], focClimSA[, "PC1"])

##### WORKING ON STUFF HERE - can't finish until SA analysis with 62 km buffer finishes #####
# x <- match(vsurfPC1$imp.mean.dec.ind, vsurfPC1_SA$imp.mean.dec.ind)
# plot(vsurfPC1$imp.mean.dec, vsurfPC1_SA$imp.mean.dec[x], type = "n")
# text(vsurfPC1$imp.mean.dec, vsurfPC1_SA$imp.mean.dec[x], 
#		 labels = climVars[vsurfPC1$imp.mean.dec.ind])
# cor.test(vsurfPC1$imp.mean.dec, vsurfPC1_SA$imp.mean.dec[x])

@@

  # polygon of confidence intervals
  x <- seq(min(focClim$bio16_cva), max(focClim$bio16_cva), l = 1e3)
  u <- predict(fitBio16cva, new = data.frame(bio16_cva = x))
  s <- sqrt(sapply(x, function(X) t(c(1, X)) %*% vcov(fitBio16cva) %*% c(1, X)))
  confInt <- data.frame(lowerCI = qnorm(0.025, u, s), 
                        upperCI = qnorm(0.975, u, s))
  polygon(c(x, rev(x)), c(confInt$lowerCI, rev(confInt$upperCI)), 
    col = "white", border = NA)
  lines(x, confInt$lowerCI, lwd = 2, lty = 2)
  lines(x, confInt$upperCI, lwd = 2, lty = 2)
  lines(range(focClim$bio16_cva), range(predict(fitBio16cva)), lwd = 2)
  points(focClim$bio16_cva, focClim$PC1, pch = 21, col = "black", 
         bg = "white", cex = 1.5, lwd = 2)
  dev.off()

@@

	plot(0, 0, type = "n", xlim = c(0, 0.2), ylim = c(1, nImpLat), 
		axes = F, frame.plot = T, xlab = "", ylab = "")
	title(xlab = "Climate-Trait PC1\nImportance", line = 4, cex.lab = 1.5)
	for (i in (nImpLat:(nImpLat - nImpPC1 + 1))) 
	{
		lines(x = c(-100, 100), y = c(i, i), col = "grey")
	}
	points(vsurfPC1$imp.mean.dec[1:nImpPC1], nImpLat:(nImpLat - nImpPC1 + 1), 
		pch = 21, col = "black", bg = "white", cex = 1.5)
	mtext(climVars[vsurfPC1$varselect.pred], side = 4, las = 1, line = 1, 
    at = nImpLat:(nImpLat - nImpPC1 + 1), 
		font = ifelse(vsurfPC1$varselect.pred %in% vsurfLat$varselect.pred,	2, 1))
	axis(1, lwd = 0, lwd.ticks = 1, at = seq(0, 0.2, 0.05))
  dev.off()

@@
